---
title: "STA5228 Reliability and Life Data Project"
subtitle: Reliability Assessment of Dishwashers. A Multimodal Failure Analysis
author: "Jezerca Hodaj"
date: "`r Sys.Date()`"
output: beamer_presentation
header-includes: 
- \usepackage{amsmath,amssymb}
- \usetheme{Madrid}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(SMRD)
# Compatibility patch for SMRD on newer versions of R
CheckPrintDataName_fixed <- function() {
  the.par <- par(xpd = TRUE)
  on.exit(par(the.par))

  if (SMRDOptions("SMRD.DateOnPlot")) {
    text(x.loc(1), y.loc(-0.2), date(), adj = 1, cex = 0.8)
  }

  text(
    x.loc(1),
    y.loc(-0.23),
    GetSMRDDefault("SMRD.NameOnPlot"),
    adj = 1,
    cex = 0.8
  )
}

environment(CheckPrintDataName_fixed) <- asNamespace("SMRD")

assignInNamespace(
  "CheckPrintDataName",
  CheckPrintDataName_fixed,
  ns = "SMRD"
)
library(MASS)
library(survival)
library(survminer)
library(ComparisonSurv)
library(muhaz)
library(rms)
library(VGAM)
library(cmprsk)
```


## Introduction

- When assessing the reliability or survival of a product very often we have to
deal with more than one failing mode. 

- Each failure mode is like a component in a series system. When one of the 
components fail the system fails.

- This project's concern is the reliability of dishwashers.

- We are going to do a parametric and non-parametric analysis on the reliability of the dishwashers of our data set. 



## Indtroduction

- Perform a parametric analysis taking into consideration each failure 
mode separately while considering the other failure mode times as right 
censured times.

- We do a non-parametric analysis, giving Kaplan-Meier estimates of the survival
 probabilities for both cases aforementioned.

- We estimate the reliability taking in consideration both failure modes. 


##  Parametric Analysis of the Dishwasher data 

- We have two failure modes, break and obstruction.

- We assume both failure modes are independent from each other.

- The failures for failure mode $j$, are marked as failures and the failures 
for the other failure mode are marked as right-censored observations.



## 2.1 Parametric Analysis for Failure Mode 1 = Break

- First we consider the failure mode break.

- Failures for failure mode break are considered as failures and the 
failures for failure mode obstruction are marked as right-censored observations.

- We need to find a distribution that fits our data. 
 For this, we use probability plots.
 
 
## 2.1 Parametric Analysis for Failure Mode 1 = Break 
 
```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=10, fig.height=6}

DishwasherBreak <- read.csv("DishwasherReliabilityBreak.csv")


## Failure Mode = Break, Censoring = Obstruction

## Weibull fits better

DishwasherBreak.ld <- frame.to.ld(DishwasherBreak,
                              response.column = 2, 
                              censor.column = 3,
                              time.units = "Cycles")


par(mfrow = c(1,2))
mleprobplot(DishwasherBreak.ld, distribution="Weibull")
legend("top", legend = "Mode 1 = Break ~ Weibull" , lty = 1:1)

mleprobplot(DishwasherBreak.ld, distribution="lognormal")
legend("top", legend = "Mode 1 = Break ~ Lognormal" , lty = 1:1)

par(mfrow = c(1,1))

```


## 2.1 Parametric Analysis for Failure Mode 1 = Break 

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=10, fig.height=6}
par(mfrow = c(1,2))
mleprobplot(DishwasherBreak.ld, distribution="loglogistic")
legend("top", legend = "Mode 1 = Break ~ Loglogistic" , lty = 1:1)

mleprobplot(DishwasherBreak.ld, distribution="frechet")
legend("top", legend = "Mode 1 = Break ~ Frechet" , lty = 1:1)
par(mfrow = c(1,1))

```


## 2.1 Parametric Analysis for Failure Mode 1 = Break 

- From these plots we see that for this failure mode Weibull distribution is a 
better fit.   

- Thus we are going to use this distribution for the parametric analysis. 

- In the next figure we have the pdf, CDF, survival and hazard function for the 
Weibull distribution with parameters, the ML estimates.


## 2.1 Parametric Analysis for Failure Mode 1 = Break 

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=5}

x<-98:2000

PDF.weibull <- dweibull(x, shape = 1.977, scale = 891.9)


q <- qweibull(x/2000,shape = 1.977, scale = 891.9)

CDF.weibull <- pweibull(q, shape = 1.977, scale = 891.9)


survival.weibull <- 1 - CDF.weibull


# Create a survival object
surv_object <- Surv(DishwasherBreak$Cycles, 
                    event = DishwasherBreak$Censoring.Indicator == "Failed" )


## Hazard function plot for the assumed distribution

hazard.fn <- psm(surv_object ~ 1, data = DishwasherBreak)



par(mfrow = c(2,2))

plot(x, PDF.weibull, type = "l", ylim = c(0, 0.001), xlab="Cycles")
plot(q, CDF.weibull, type = "l",  xlab="Cycles")
plot(q, survival.weibull, type = "l",  xlab="Cycles")
survplot(hazard.fn, what="hazard")

par(mfrow = c(1,1))


```
 

## 2.1 Parametric Analysis for Failure Mode 1 = Break 

- Under the assumption that the Weibull distribution is correct, we have 

- The ML estimate for the Mean Time To Failure is $MTTF = 790.6$ cycles

- We can see that $10\%$ of the dishwashers will fail due to arm spray break 
after about 87 cycles

- and $90\%$ of them will fail after about 1360 cycles 



## 2.1 Parametric Analysis for Failure Mode 1 = Break 

- After 1000 cycles the probability that the spray arm will break is $0.7145$,
with a $95\%$ confidence interval $(0.5712, 0.8437)$.

- The hazard rate is very small, which tells us that in general the spray arms
take longer to break, after having completed many cycles. 

- ML Estimates table

```{r,message=FALSE, warning= FALSE, echo=FALSE}

DishwasherBreak.mlest <- mlest(DishwasherBreak.ld, 
                             distribution = "Weibull")

round(DishwasherBreak.mlest$mle.table,4)


```



## 2.1 Parametric Analysis for Failure Mode 1 = Break 

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=5, fig.height=5}
DishwasherBreak.mlest$failure.probabilities
```

## 2.1 Parametric Analysis for Failure Mode 1 = Break 

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=4, fig.height=5}
DishwasherBreak.mlest$quantiles
```



## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction


- Now we consider the failure mode obstruction.

- Failures for failure mode obstruction as failures and the failures for 
failure mode break are marked as right-censored observations. 

- For the parametric analysis we need to find a distribution that fits well our 
data.

- Next slide we have the probability plots for four different distributions.


## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=8, fig.height=7}


DishwasherObstruction <- read.csv("DishwasherReliabilityObstruction.csv")


## Failure Mode = Break, Censoring = Obstruction

##Lognormal fits better


DishwasherObstruction.ld <- frame.to.ld(DishwasherObstruction,
                                        response.column = 2, 
                                        censor.column = 3,
                                        time.units = "Cycles")


par(mfrow= c(1,2))
mleprobplot(DishwasherObstruction.ld, distribution="Weibull")
legend("top", legend = "Mode 2 = Obstruction ~ Weibull" , lty = 1:1)

mleprobplot(DishwasherObstruction.ld, distribution="lognormal")
legend("top", legend = "Mode 2 = Obstruction ~ Lognormal" , lty = 1:1)
par(mfrow= c(1,1))
```

## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=8, fig.height=7}

par(mfrow= c(1,2))
mleprobplot(DishwasherObstruction.ld, distribution="loglogistic")
legend("top", legend = "Mode 2 = Obstruction ~ Loglogistic" , lty = 1:1)

mleprobplot(DishwasherObstruction.ld, distribution="frechet")
legend("top", legend = "Mode 2 = Obstruction ~ Frechet" , lty = 1:1)
par(mfrow= c(1,1))

```


## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction

- From these four plots, Frechet distributions seems to be a better fit for
the failure mode obstruction. 

- In the next slide we have the pdf, CDF, survival and hazard function for
the Frechet distribution with parameters, the ML estimates.



## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=7, fig.height=6}

x<-7:1200

PDF.frechet <- dfrechet(x, scale = 1.95, shape = 4.928)


q <- qfrechet(x/1200,scale = 1.95, shape = 4.928)

CDF.frechet <- pfrechet(q, scale = 1.95, shape = 4.928)

survival.frechet <- 1-CDF.frechet


# Create a survival object
surv_object1 <- Surv(DishwasherObstruction$Cycles, 
              event = DishwasherObstruction$Censoring.Indicator == "Failed")

## Hazard function plot for the assumed distribution

hazard.fn1 <- psm(surv_object1 ~ 1, data = DishwasherObstruction)


par(mfrow = c(2,2))

plot(x, PDF.frechet, type = "l", ylim = c(0, 0.0015), xlab="Cycles")
plot(q, CDF.frechet, type = "l",  xlab="Cycles")
plot(q, survival.frechet, type = "l",  xlab="Cycles")
survplot(hazard.fn1, what="hazard")

par(mfrow = c(1,1))

```



## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction


 - The ML estimate for the Mean Time To Failure under this distribution does not
 exist, since the shape parameter $\beta = \frac{1}{\sigma} = 0.513 < 1$.
 
 - $10\%$ of the dishwashers will fail due to arm spray obstruction after about 
 7 cycles
 
 - $90\%$ of them will fail after about 11120 cycles.
 
 - After 1000 cycles the probability that the spray arm will break is $0.6961$,
 with a $95\%$ confidence interval $(0.5733, 0.7898)$. 
 
 - The hazard rate is decreasing, with the increase in the number of cycles. 
 
 
## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction



```{r,message=FALSE, warning= FALSE, echo=FALSE}

DishwasherObstruction.mlest <- mlest(DishwasherObstruction.ld, 
                               distribution = "frechet")


round(DishwasherObstruction.mlest$mle.table, 4)

```

## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=4, fig.height=5}

DishwasherObstruction.mlest$failure.probabilities

```


## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=4, fig.height=5}

DishwasherObstruction.mlest$quantiles

```


## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction

- We construct a multiple probability plot that shows the parametric and 
non-parametric estimates for each of the failure modes together in one plot.


## 2.2 Parametric Analysis for Failure Mode 2 = Obstruction

```{r,message=FALSE, warning= FALSE, echo=FALSE}

mleprobplot(DishwasherBreak.ld, distribution="Weibull", lwd.ci = NULL)

mleprobplot(DishwasherObstruction.ld, distribution="frechet", 
            param.loc = "topleft", lwd.ci = NULL, add = TRUE)


legend("top", legend = c("Mode 2 = Obstruction ~ Frechet",
                         "Mode 1 = Break ~ Weibull" ) , lty = 1:1)

```





## 3. Non-Parametric Analysis

- Kaplan-Meier Estimators for both failure modes 


## 3.1 Non-Parametric Analysis for Failure Mode 1 = Break

- We are going to compute the Kaplan-Meier estimates for the failing and survival probabilities. 

- The table below gives the Kaplan-Meier Estimates


## 3.1 Non-Parametric Analysis for Failure Mode 1 = Break

- The top part of the table for the Kaplan-Meier estimates

```{r,message=FALSE, warning= FALSE, echo=FALSE}
DishwasherBreak1 <- read.csv("DishwasherReliabilityBreak2.csv")


ti <- DishwasherBreak1$Cycles
di <- DishwasherBreak1$Failed
ri <- 0
n <- sum(DishwasherBreak1$Failed)
ni <- DishwasherBreak1$Risked
pi.hat<- di/ni
S.hat<- cumprod(1-pi.hat)
F.hat <- 1 - S.hat

DishwasherBreak.1 <- cbind(DishwasherBreak1, DishwasherBreak1$Cycles, S.hat, F.hat)

# Filter rows where Failure is "Break" and select S.hat and F.hat
result <- DishwasherBreak.1[DishwasherBreak1$Failure == "Break", c("Failure","Cycles","S.hat", "F.hat")]

# View the resulting data frame
print(head(result))


```

## 3.1 Non-Parametric Analysis for Failure Mode 1 = Break

- Here we have the bottom part of the table

```{r,message=FALSE, warning= FALSE, echo=FALSE}
print(tail(result))


```



## 3.1 Non-Parametric Analysis for Failure Mode 1 = Break

- We construct a probability plot for the failures of failure mode break.

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}

subset_break <- DishwasherBreak.1[DishwasherBreak1$Failure == "Break", ]

lab.vec <- c(0.0001,0.0002,0.0003,0.0005,0.001,0.003,0.005,0.01,0.02,0.03,
             0.05,0.1,0.2,0.3, 0.5, 0.8, 0.9, 0.99, 0.999)


plot(subset_break$Cycles, log(-log(1 - subset_break$F.hat)),
     ylab = "Probability", xlab = "Cycles",
     main = "Weibull Probability Scale",
     ylim = log(-log(1 - c(0.001, 0.999))), log = 'x', axes = FALSE,
     xlim = c(98, 2000))

# Add axes
axis(1)
axis(2, las = 2, at = log(-log(1 - lab.vec)), tcl = 0.1, labels = lab.vec)

```



## 3.1 Non-Parametric Analysis for Failure Mode 1 = Break

- Here we use the survival function to estimate the survival probabilities for each cycle and the corresponding $95\%$ confidence intervals. 

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}
# Create a survival object
surv_object <- Surv(DishwasherBreak1$Cycles, 
                    event = DishwasherBreak1$Censoring.Indicator == "Failed")


survfit_object <- survfit(surv_object ~ 1, data = DishwasherBreak1)

summary_result<- summary(survfit_object, time = DishwasherBreak1$Cycles)



# Create a data frame
summary_df <- data.frame(time = summary_result$time, 
                         survival = summary_result$surv,
                         n.risk = summary_result$n.risk,
                         n.event = summary_result$n.event,
                         std.err = summary_result$std.err,
                         "lower 95 CI" = summary_result$lower,
                         "upper 95 CI" = summary_result$upper)


summary_df_break <- subset(summary_df, n.event == 1)

table_summary <- cbind(Cycles = summary_df_break$time, Survival=summary_df_break$survival, Std.err=summary_df_break$std.err, "Lower 95%"=summary_df_break$lower.95.CI, "Upper 95%"=summary_df_break$upper.95.CI)


head(table_summary)

```


## 3.1 Non-Parametric Analysis for Failure Mode 1 = Break


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}

tail(table_summary)

```

- The two different ways of computing the survival probabilities give similar results

## 3.1 Non-Parametric Analysis for Failure Mode 1 = Break

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}

# Plot the Kaplan-Meier survival curve

# Create a survival object
surv_object.1 <- Surv(subset_break$Cycles, 
                    event = subset_break$Censoring.Indicator == "Failed")


survfit_object.1 <- survfit(surv_object.1 ~ 1, data = subset_break)


plot(survfit_object.1, col = 1, mark.time = TRUE, main = "Kaplan-Meier Survival Curve",
     xlab = "Time to Event", ylab = "Survival Probability")



```



## 3.2 Non-Parametric Analysis for Failure Mode 2 = Obstruction

- We repeat the same process as for failure mode 1 = break. 

- Compute the Kaplan-Meier estimators, and construct a probability plot for 
obstruction failure times. 



## 3.2 Non-Parametric Analysis for Failure Mode 2 = Obstruction

- Top part of the table for the Kaplan-Meier estimates

```{r,message=FALSE, warning= FALSE, echo=FALSE}


DishwasherObstruction1 <- read.csv("DishwasherReliabilityObstruction2.csv")

#plot the hazard cumulative function for failure mode = break

ti <- DishwasherObstruction1$Cycles
di <- DishwasherObstruction1$Failed
n <- sum(DishwasherObstruction1$Failed)
ni <- DishwasherObstruction1$Risked
pi.hat<- di/ni
S.hat1<- cumprod(1-pi.hat)
F.hat1 <- 1 - S.hat1


DishwasherObstruction.1 <- cbind(DishwasherObstruction1,DishwasherObstruction1$Cycles, S.hat1, F.hat1)

# Filter rows where Failure is "Break" and select S.hat and F.hat
result.o <- DishwasherObstruction.1[DishwasherObstruction1$Failure == "Obstruction", 
c("Failure","Cycles","S.hat1", "F.hat1")]

# View the resulting data frame
print(head(result.o))

```

## 3.2 Non-Parametric Analysis for Failure Mode 2 = Obstruction

- Bottom part of the table

```{r,message=FALSE, warning= FALSE, echo=FALSE}

print(head(result.o))

```


## 3.2 Non-Parametric Analysis for Failure Mode 2 = Obstruction


- Probability plot for the failures of failure mode obstruction. 


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}

subset_obst <- DishwasherObstruction.1[DishwasherObstruction1$Failure == "Obstruction", ]

lab.vec <- c(0.0001,0.0002,0.0003,0.0005,0.001,0.003,0.005,0.01,0.02,0.03,
             0.05,0.1,0.2,0.3, 0.5, 0.8, 0.9, 0.99, 0.999)


plot(subset_obst$Cycles, log(-log(1 - subset_obst$F.hat1)),
     ylab = "Probability", xlab = "Cycles",
     main = "Weibull Probability Scale",
     ylim = log(-log(1 - c(0.001, 0.999))), log = 'x', axes = FALSE,
     xlim = c(7, 1500))

# Add axes
axis(1)
axis(2, las = 2, at = log(-log(1 - lab.vec)), tcl = 0.1, labels = lab.vec)


```


## 3.2 Non-Parametric Analysis for Failure Mode 2 = Obstruction

- Using the survival function to estimate the survival probabilities for each cycle and the corresponding $95\%$ confidence intervals. 


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}
# Create a survival object
surv_object1 <- Surv(DishwasherObstruction1$Cycles, 
              event = DishwasherObstruction1$Censoring.Indicator == "Failed")


survfit_object1 <- survfit(surv_object1 ~ 1, data = DishwasherObstruction1)

summary_result1<- summary(survfit_object1, time = DishwasherObstruction1$Cycles)

# Create a data frame
summary_df1 <- data.frame(time = summary_result1$time, 
                         survival = summary_result1$surv,
                         n.risk = summary_result1$n.risk,
                         n.event = summary_result1$n.event,
                         std.err = summary_result1$std.err,
                         "lower 95 CI" = summary_result1$lower,
                         "upper 95 CI" = summary_result1$upper)


summary_df_obst <- subset(summary_df1, n.event == 1)


table_summary1 <- cbind(Cycles = summary_df_obst$time, Survival=summary_df_obst$survival, Std.err=summary_df_obst$std.err, "Lower 95%"=summary_df_obst$lower.95.CI, "Upper 95%"=summary_df_obst$upper.95.CI)


head(table_summary1)

```



## 3.2 Non-Parametric Analysis for Failure Mode 2 = Obstruction


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}

tail(table_summary1)

```


- The two different ways of computing the survival probabilities give similar results



## 3.2 Non-Parametric Analysis for Failure Mode 2 = Obstruction


```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}

# Create a survival object
surv_object.2 <- Surv(subset_obst$Cycles, 
                      event = subset_obst$Censoring.Indicator == "Failed")


survfit_object.2 <- survfit(surv_object.2 ~ 1, data = subset_obst)



# Plot the Kaplan-Meier survival curve

plot(survfit_object.2, col = 1, mark.time = TRUE, main = "Kaplan-Meier Survival Curve",
     xlab = "Time to Event", ylab = "Survival Probability")


```



## 4. Consider both Failure Modes

- Let $T_{1}$ and $T_{2}$ be the failure times for failure mode 1 and failure mode 2 respectively.

- Then, the joint distribution of $T_{1}$ and $T_{2}$ is $F(t_{1}, t_{2}) = \Pr (T_{1} \le t_{1}, T_{2} \le t_{2})$ and it completely describes the multiple failure modes
probability model. 

- The marginal distribution for mode $j$ is $F_{j}(t) = \Pr(T_{j} \le t)$.

- We can interpret this as the cdf for mode $j$ if only mode $j$ is active, 
where $j = 1, 2$ in our case.

- We assume that failure modes are independent from each other. 


## 4. Consider both Failure Modes

- The joint distribution of the two failure modes is 

\[
F(t_{1}, t_{2}) = \prod_{j = 1}^J F_{j}(t_{j})
\]

- The system failure time is the first failure time. 

- 
\[
T = \min(T_{1}, T_{2})
\] 

- \[
F_{T}(t) = \Pr(T \le t) = 1 - \Pr(T > t) = 1 - \Pr(T_{1} >t_{1} \cap T_{2} > t_{2} )
\]

\[
F_{T}(t) = 1 - \prod_{j=1}^{2}[1 - F_{j}(t)]
\]


## 4. Consider both Failure Mode

- For failure mode 1 = break we assumed that the failures followed a Weibull
distribution

- For failure mode 2 = obstruction, we assumed a Frechet distribution. 

- The system's failure-time cdf 

- \[
F_{T}(t) = 1 - \exp \Big[ - (\frac{t}{\eta_{1}})^{\beta_{1}}\Big] + \exp \Bigg[ - (\frac{t}{\eta_{1}})^{\beta_{1}} - (\frac{t}{\eta_{2}})^{-\beta_{2}} \Bigg]
\] 

- where, $\eta_{2} = \exp(\mu)$ and $\beta_{2} = \frac{1}{\sigma}$.

## 4. Consider both Failure Mode

- An estimate of the failure-time cdf for the system with all failure modes 
active is obtained by substituting the ML estimates for the parameters for 
all of the failure modes into the system cdf. 

- Since we already obtained the ML estimates for all the failure modes, 
we can estimate the cdf of the system failure-time.

- The ML estimates are :

## 4. Consider both Failure Mode

```{r,message=FALSE, warning= FALSE, echo=FALSE}

Dishwasher2 <- read.csv("DishwasherReliability2.csv")

eta.w <- 891.9
beta.w <- 1.977

mu.f <- 4.928
sigma.f <- 1.95

eta.f <- exp(mu.f)
beta.f <- 1/sigma.f


F.hat.joint <- 1 - exp(-1*(Dishwasher2$Cycles/eta.w)^beta.w) + 
  exp(-1*((Dishwasher2$Cycles/eta.w)^beta.w + (eta.f/Dishwasher2$Cycles)^beta.f))

S.hat.joint <- 1- F.hat.joint

Cycles <- Dishwasher2$Cycles

head(cbind(Cycles, F.hat.joint, S.hat.joint))

```



## 4. Consider both Failure Mode

```{r,message=FALSE, warning= FALSE, echo=FALSE}
tail(cbind(Cycles, F.hat.joint, S.hat.joint))

```


## 4. Consider both Failure Mode

- Kaplan-Meier estimates for system failure-times

## 4. Consider both Failure Mode

- Top part of the table for the Kaplan-Meier estimates

```{r,message=FALSE, warning= FALSE, echo=FALSE}

Dishwasher2 <- read.csv("DishwasherReliability2.csv")
# Kaplan - Meier Estimate 

ti <- Dishwasher2$Cycles
di <- Dishwasher2$Failed
n <- sum(Dishwasher2$Failed)
ni <- Dishwasher2$Risked
pi.hat<- di/ni
S.hat.KM<- cumprod(1-pi.hat)
F.hat.KM <- 1 - S.hat.KM

Cycles <- Dishwasher2$Cycles

Dishwasher.2 <-cbind(Cycles,F.hat.KM, S.hat.KM)
print(head(Dishwasher.2))

```

## 4. Consider both Failure Mode

- Bottom part of the table for the Kaplan-Meier estimates

```{r,message=FALSE, warning= FALSE, echo=FALSE}

print(tail(Dishwasher.2))


```


## 4. Consider both Failure Mode

- Weibull Probability Plot Scale for proportion failing

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}
### Plotting the Kaplan - Meier estimate 

lab.vec <- c(0.0001,0.0002,0.0003,0.0005,0.001,0.003,0.005,0.01,0.02,0.03,
             0.05,0.1,0.2,0.3, 0.5, 0.8, 0.9, 0.99, 0.999)


plot(Cycles, log(-log(1 - F.hat.KM)),
     ylab = "Probability", xlab = "Cycles",
     main = "Weibull Probability Scale",
     ylim = log(-log(1 - c(0.001, 0.999))), log = 'x', axes = FALSE,
     xlim = c(7, 2000))

# Add axes
axis(1)
axis(2, las = 2, at = log(-log(1 - lab.vec)), tcl = 0.1, labels = lab.vec)


```



## 5. Comparing the two failure mode distributions


- Previously we assumed that each of the two failure modes followed two 
different distributions, Weibull and Frechet, respectively.

- Let's suppose that also the second failure mode, obstruction, follows a 
Weibull Distribution.

- We can perform a Likelihood ratio test, to check if under this new assumption,
there is any statistically significant difference between the two failure modes


## 5. Comparing the two failure mode distributions

- Parameter estimates and log-likelihood for failure mode Break

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}
Dishwasher <- read.csv("DishwasherReliability1.csv")


surv_object3 <- Surv( time = Dishwasher$Cycles
                         ,time2 = Dishwasher$Cycles
                         ,type = "interval2")


#Fitting Separate Distributions to each failure mode
SepDist.fits <- lapply(c("Break", "Obstruction"),
                       function(failure)
                         survreg(Surv( time = Cycles
                                       ,time2 = Cycles
                                       ,type = "interval2") ~ 1,
      data=subset(Dishwasher, Dishwasher$Failure == failure),
                                 dist = "weibull"))


print(SepDist.fits[[1]])

```


## 5. Comparing the two failure mode distributions

- Parameter estimates and log-likelihood for failure mode Obstruction

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}

print(SepDist.fits[[2]])

log.l.break <- -184.1

log.l.obstruction <- -243.9 

log.l.sep <- log.l.break + log.l.obstruction

```


## 5. Comparing the two failure mode distributions

- For the marginal distribution of failure mode break, we have that the maximum 
likelihood is -184.1, 

- For the marginal distribution of failure mode obstruction the maximum 
likelihood is -243.9. 

- The sum of the two equals -428. 

- Next we fit distributions with the same sigma parameter for both.



## 5. Comparing the two failure mode distributions

```{r,message=FALSE, warning= FALSE, echo=FALSE, fig.width=6, fig.height=4}

#Assuming both failure modes have equal sigma

Equal.fit <- survreg(surv_object3 ~ as.factor(Failure),
                     data = Dishwasher,
                     dist="weibull")

print(Equal.fit)

log.l.equal <- -433.1

LRtest <- 2*(log.l.sep -log.l.equal)

c.value <- round(qchisq(0.95, 1),3)

```



## 5. Comparing the two failure mode distributions

- Equal sigma distributions gives a maximum likelihood equal to
 -433.1.

- Then the LRtest = 10.2, whereas $chisq(0.95, 1) = 3.841$. 

- Since LRtests > $chisq(0.95, 1)$, then we reject the hypothesis that both distributions have equal sigmas.

- Therefore, the two failure modes have statistically significant differences between each-other.



## 6. Conclusions 

- The Mean Time To Failure of spray arms due to breaks was 790.6 cycles. 

- For obstructions the MTTF did not exist under the Frechet, distribution. 

- From Kaplan-Meier estimators: 

- About 95% of the spray arms survived breaks for at least 141.9 cycles

- 95% of spray arms survived obstruction for at least 10.2 cycles.


## 6. Conclusions 

- About 95% of spray arms survived both breaks and obstructions for at 
least 10.2 cycles. 

- From our results it seems that the obstruction failure mode needs more 
attention from the dishwasher engineers. 

- They should focus on finding solutions on how to make the spray arm more 
resistant to obstructions. 


## 7. The End

                            Thank you!

